To my mother Yaxian and my father Jun, who have provided me love and emotional support with understanding and patience.

I would like to thank my advisor Professor Nicolas Schweighofer, who has provided guidance and swept away many blocks in the route of research. 
Also to my committee member Professor James Gordon and Professor Stefan Schaal, who evaluated my work with a critical eye and always gave valuable comments and feedback. 
Also to Professor Finley and Professor Liew who, together with all other members of Computational Motor Control and Learning Journal Club, provided a friendly environment for discussion.

Thanks to all my labmates: Yupeng, Sujin, Amar, Hyeshin, Youngmin, Victor and others in CNRL lab; BK, Irene, Andrew, Dorsa, Clarisa, Rini, Yuchen, Yi'an and others in MBNL lab; Chang, Natalia, Aram, Sungwoo and others in LCL lab.
Many cheerful (or tiresome) and fruitful discussions with them that brought insight to my work, and many small talks and chitchats in which we shared our gossips and complaints about our advisors and TA works.

Grateful to all program staff: Dawn Burke, Ariana Perez and Deanna Solorzano in NGP; Veronica Perez and Lydia Vazquez in BKN for their unfailing support and assistance.

Special thanks to my friend Zixuan, whose vibe and grit encourage me not to fall too behind;
To Tian Chen, who boosted my confidence at times of difficulties; 
To my friend Liu Chang, whose cheerfulness was highly infective.
To my friend Zhiqiang, Zhihui, Feng Ying and others in the Youth Group whose altruistic and cheerful characters remind me the world is still good.
Thanks to my fellow badminton players Wenjian and Tianjiao, together with whom I forget everything in the world enjoying the sport.
And to all my friends these years along the way who listened to my boring problems and provided support in many ways.



